---
Title: "Simulation of an Exponential Distribution and the Central Limit Theorem"
Author: "Manuel A. Cerda R."
Date: "31 de July de 2016"
Output: html_document
---

#Overview

This simulation investigates the exponential distribution in R and compares it with the Central Limit Theorem. The exponential distribution is simulated in R wirh rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda where lambda = 0.2 for all of the simulations. The distribution of averages is based upon 40 exponentials. And, a total of a 1000 simulations are considered.


###Content

1. Simulations
2. Sample Mean versus Theoretical Mean
3. Sample Variance versus Theoretical Variance
4. Distributions


###Simulations

```{r sampling, echo=TRUE}
library(knitr)
library(ggplot2)

# setting the random generator seed to ensure replicable steps
set.seed(1234);

# declaring known variables 
# (lambda, theoretical mean, standard deviation, number of distribution averages, simulations)
lambda <- 0.2
t_mean <- 1/lambda
sd <- 1/lambda
n <- 40
n_sim <- 1000

# generating n=40 random exponentials
my_rexp <- function() { rexp(n, rate = lambda) }
summary(my_rexp())

# generating the 1000 simulated trials
simTrials <- data.frame(values=replicate(n_sim, mean(my_rexp() )))
summary(simTrials$values)

# calculating cumputable variables
# (theoretical standard deviation)
t_sd <- sd * 1 / sqrt(n)
```


###Sample Mean versus Theoretical Mean

```{r smean_tmean, echo=TRUE}
# Random Exponential Distribution
hist(simTrials$values, freq= T, ylab= "Frequency", xlab="Random Exponential", main="Random Exponential Distribution", col=rgb(0, 1, 0, 0.25))
abline(v = t_mean, col="red", lwd = 2, lty = 1)

```


###Sample Variance versus Theoretical Variance



###Distributions